Statically Typed String Sanitation Inside a Python

### 2014-08-06 sanitation-psp14/sanitation-psp14-short.tex
Changes to soft stuff

%\category{D.3.2}{Programming Languages}{Language Classifications}[Extensible Languages]
%\category{D.3.4}{Programming Languages}{Processors}[Compilers]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
%\keywords
%type-level computation, typed compilation
\section{Introduction}\label{intro}
Improper input sanitation is a leading cause of security vulnerabilities in web applications [OWASP]. Command injection attacks exploit improper input sanitation by inserting malicious code into an otherwise benign command. Modern web frameworks, libraries, and database abstraction layers attempt to ensure proper sanitation of user input. When these methods are unavailable or insufficient, developers implement custom sanitation techniques. In both cases, sanitation algorithms are implemented using the language's regular expression capabilities and usually \emph{replace} potentially unsafe strings with equivalent escaped strings.

@@

A number of research languages provide static guarantees that a program is free of input sanitation vulnerabilities [Jif][Ur/Web]. 
Unlike this work, our solution to the input sanitation problem has a very low barrier to adoption; for instance, our implementation conservatively extends Python -- a popular language among web developers.
We also believe our general approach is better-positioned for security, where continuously evolving threats might require frequent addition of new analyses; in these cases, the composability and generality of our approach is a substantial advantage.

@@

We are also unaware of any extensible programming languages which emphasize applications to security concerns (TRUE?).

Incorporating regular expressions into the type system is not novel. The XDuce system \cite{pierce} typechecks XML schemas using regular expressions. We differ from this and related work in at least two ways. First, our system  is defined within an extensible type system; second, and more importantly, we have demonstrated that regular expression types are applicable to the web security domain.

@@

In this section we define a language for statically checked string sanitation ($\lambdas$)
and prove that its correctness property is preserved under translation to a language
with regular expression matching capabilities ($\lambdap$). A brief outline of this section
follows:
\begin{itemize}
  \item Page 3 contains a definition of $\lambdas, \lambdap$ and the translation from $\lambdas$ to $\lambdap$.
  \item In \S 2.1 we state some properties about regular expressions which are needed in our correctness proofs.
  \item In \S 2.2 we prove type safety for $\lambdap$ as well as both type safety and correctness for $\lambdas$.
  \item In \S 2.3 we prove that translation preserves the correctness reesult about $\lambdas$.
\end{itemize}

@@

Our type safety proof for language S replies on a relationship between
string substitution and language substitution given in lemma \ref{thm:substcorrespondence}.
We also rely upon several other properties of regular languages.
Throughout this section, we fix an alphabet $\Sigma$ over which strings $s$ and
regular expressions $r$ are defined. throughout the paper, $\lang{r}$ refers to the
language recognized by the expression $r$. This distinction between the expression
and its language -- typically elided in the literature -- makes our definition and
proofs about systems S and P more readable.

\begin{lem}{Properties of Regular Languages and Expressions.} \label{thm:regexprops}
The following are properties of regular expressions which are necessary for our proofs:
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $s_1s_2 \in \lang{r_1r_2}$.
For all strings $s$ and expressions $r$, either $s \in \lang{r}$ or $s \not \in \lang{r}$.
\end{lem}

@@

\begin{lem}{Closure.} \label{thm:total}
  If $\lang{r}, \lang{r_1}$ and $\lang{r_2}$ are regular expressions, then $\lang{\lreplace{r}{r_1}{r_2}}$ is also a regular language.
\end{lem}
\begin{proof}
The theorem follows from closure under difference, right quotient and reversal.
\end{proof}

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the refinitions of lsubst and lreplace; note that language substitutions over-approximate string substititons.
\end{proof}

@@

In addition to safety, we proof a correctness result for $\lambdas$ which will
be preserved under translation.

\begin{thm}{Correctness of Input SAnitation for $\lambdas$.}\label{thm:scorrect}
  If  $\sctx \vdash S : \stringin{r}$ and $S \sreduces \rstr{s}$ then $s \in \lang{r}$.
\end{thm}
\begin{proof}
  Follows directly from type safety, canonical forms for $\lambdas$.
\end{proof}

@@

@@

Papers that needs to be cited in this section:
\begin{itemize}
  \item Ur/Web OSDI paper
  \item Jif?
  \item OWASP
  \item XDuce and related papers. 
  \item src?
  \item Ace or Wyvern paper?
  \item hotsos?
  \item Haskell extension paper
  \item Maybe some popular FOSS libraries/frameworks that do input sanitation?
\end{itemize}

@@

%\bibliographystyle{abbrv}
%% The bibliography should be embedded for final submission.
%\bibliography{../research}
%%\softraggedright
%%P. Q. Smith, and X. Y. Jones. ...reference text...

\end{document}

>>>>

%\category{D.3.2}{Programming Languages}{Language Classifications}[Extensible Languages]
%\category{D.3.4}{Programming Languages}{Processors}[Compilers]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
%\keywords
%type-level computation, typed compilation
\section{Introduction}\label{intro}
Improper input sanitation is a leading cause of security vulnerabilities in web applications \cite{OWASP}. Command injection attacks exploit improper input sanitation by inserting malicious code into an otherwise benign command. Modern web frameworks, libraries, and database abstraction layers attempt to ensure proper sanitation of user input. When these methods are unavailable or insufficient, developers implement custom sanitation techniques. In both cases, sanitation algorithms are implemented using the language's regular expression capabilities and usually \emph{replace} potentially unsafe strings with equivalent escaped strings.

@@

A number of research languages provide static guarantees that a program is free of input sanitation vulnerabilities \cite{UrFlowOSDI10}. 
Unlike this work, our solution to the input sanitation problem has a very low barrier to adoption; for instance, our implementation conservatively extends Python -- a popular language among web developers.
We also believe our general approach is better-positioned for security, where continuously evolving threats might require frequent addition of new analyses; in these cases, the composability and generality of our approach is a substantial advantage.

@@

We are also unaware of any extensible programming languages which emphasize applications to security concerns.

Incorporating regular expressions into the type system is not novel. The XDuce system \cite{HosoyaVouillonPierce2000ICFP,HosoyaPierce2002} checks XML documents against schemas using regular expressions. 
Similarly, XHaskell \cite{xhaskell} focuses on XML documents.
We differ from this and related work in at least two ways. First, our system  is defined within an extensible type system; second, and more importantly, we have demonstrated that regular expression types are applicable to the web security domain.

@@

In this section we define a language for statically checked string sanitation ($\lambdas$).
The system has regular expression types $\stringin{r}$ whre $r$ is a regular expression. 
Expressions of this type evaluate to string literals in the language described by $r$.
Operations on expressions of type $\stringin{r}$ preserve this property; for instance,
the result of string concatenation is

The premier operation for manipulating strings in $\lambdas$ is string substitution, 
which is a familiar operation to any programmer who has used regular expressions.
The replacement operation replaces all instances of a pattern in one string with
another string; for instance, $\lsubst(a|b, ``a", ``"c") = c$.
In order to computer the type resulting from aubstitution, we also need to compute
the result of replacing on language with another inside a given language.
Finally, just for convienance, we provide a coerce operation. The introduction of
coercion requires handling of runtime errors.

The underlying language $\lambdap$ has only one type for strings. We prove that whenever
a term is translated from $\lambdas$ to $\lambdap$, correctness is preserved.
The only exception is in the case of unsafe casts in $\lambdas$, which are unnecessary but
are included to demonstrate that the regex library of $\lambdap$ may be used to insert
dynamic checks whenever even when developers are not careful about using statically checked
operations.

A brief outline of this section follows:
\begin{itemize}
  \item Page 3 contains a definition of $\lambdas, \lambdap$ and the translation from $\lambdas$ to $\lambdap$.
  \item In \S 2.1 we state some properties about regular expressions which are needed in our correctness proofs.
   \item In \S 2.2 we prove type safety for $\lambdap$ as well as both type safety and correctness for $\lambdas$.
  \item In \S 2.3 we prove that translation preserves the correctness reesult about $\lambdas$.
\end{itemize}

@@

Our type safety proof for language S replies on a relationship between
string substitution and language substitution given in lemma \ref{thm:substcorrespondence}.
We also rely upon several other properties of regular languages.
Throughout this section, we fix an alphabet $\Sigma$ over which strings $s$ and
regular expressions $r$ are defined. throughout the paper, $\lang{r}$ refers to the
language recognized by the regular expression $r$. This distinction between the regular expression
and its language -- typically elided in the literature -- makes our definition and
proofs about systems S and P more readable.

\begin{lem}{Properties of Regular Languages and Expressions.} \label{thm:regexprops}
The following are properties of regular expressions which are necessary for our proofs:
If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $s_1s_2 \in \lang{r_1r_2}$.
For all strings $s$ and regular expressions $r$, either $s \in \lang{r}$ or $s \not \in \lang{r}$.
Regular languages are closed under difference, right quotient, reversal, and string homomorphism.
\end{lem}

If any of these properties are unfamiliar, the reader may refer to a standard text on the subject \cite{cinderella}.

@@

\begin{lem}{Closure.} \label{thm:total}
  If $\lang{r}, \lang{r_1}$ and $\lang{r_2}$ are regular expressions, then $\lang{\lreplace{r}{r_1}{r_2}}$ is also a regular language.
\end{lem}
\begin{proof}
The theorem follows from closure under difference, right quotient, reversal and string homomorphism.
\end{proof}

\begin{lem}{Substitution Correspondence.} \label{thm:substcorrespondence}
  If $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ then $\lsubst{r}{s_1}{s_2} \in \lang{\lreplace{r}{s_1}{s_2}}$.
\end{lem}
\begin{proof}
The theorem follows from the definitions of lsubst and lreplace; note that language substitutions over-approximate string substititons.
\end{proof}

@@

In addition to safety, we proof a correctness result for $\lambdas$ which ensures that well-typed terms of regular string type are in the language associated with their type.

\begin{thm}{Correctness of Input Sanitation for $\lambdas$.}\label{thm:scorrect}
  If  $\sctx \vdash S : \stringin{r}$ and $S \sreduces \rstr{s}$ then $s \in \lang{r}$.
\end{thm}
\begin{proof}
  Follows directly from type safety, canonical forms for $\lambdas$.
\end{proof}

@@

\section{Implementation}

An implementation of a similar system as a type system extension in the programming language Ace is discussed in \cite{fulton12} and \cite{fulton13}.
The implementation only supports a special case of replacement where unsafe substrings are simply stripped.
In practice, most libraries and frameworks escape command sequences instead of stripping characters; therefore, we plan to extend our implementation with
support for the replace operation as described in this paper..

@@

@@

\bibliographystyle{abbrv}

\bibliography{research}

% The bibliography should be embedded for final submission.

%\begin{thebibliography}
%
%\bibitem{lamport94}
%Leslie Lamport,
%\emph{\LaTeX: a document preparation system}.
%Addison Wesley, Massachusetts,
%2nd edition,
%1994.
%
%\end{thebibliography}
\end{document}
